---
title: "Transportation Network Companies (TNCs) Effect on Rail Ridership"
author: "Charlie Berman"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("../code/housekeeping.R")

```

## Background and Literature
* Declining Public Transportation Usage
* First Mile/Last Mile Problem (FM/LM)
* TNC's Impact on Public Transportation
  * Reduces public transit demand in cities with "high urban compactness" (Lee, et. al 2019)
  * Serves as compliment (net) but masks heterogeneity (Hall, et. al 2018)
  * Serves as compliment to high quality public transit but substitute to low quality transit (Zhao 2019)
  
## Research Question and Methodology

* Can Uber be used to supplement public transportation by offering a solution to the FM/LM problem?
  * Diff-in-diff comparing Buffalo and similar cities
  * Spatial analysis of customers added with introduction of Uber (?)
  
```{r,echo=FALSE, message=FALSE, out.width="70%",fig.align='center'}
df <- read_csv("../data/buff_clev_simple.csv")
df <- df %>%
  filter(year < 2020)
ggplot(df, aes(x = date, y = ridership, group = agency, color = as.factor(agency))) +
  geom_line() +
  labs(title = "Ridership Over Time",
       x = "Date",
       y = "Ridership",
       color = "Agency") +
  scale_y_continuous(labels = scales::number_format()) +
  geom_vline(xintercept = as.Date("2014-04-01"), color = "blue", linetype = "solid") + # Add a red dashed vertical line
  geom_vline(xintercept = as.Date("2014-03-13"), color = "green", linetype = "dashed") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Optional: Rotate x-axis labels for better visibility
```


## Data Sources
* National Transit Database (NTD)
  * Contains monthly ridership data from 2002-present for all public transportation agencies that receive federal funding
* Google search data

```{r, echo=FALSE, message=FALSE, out.width="70%"}
both_rows <- rbind(head(df, n=3), tail(df, n=3))
both_rows %>%
  kable("html", align = "c") %>%
  kable_styling(full_width = FALSE, position = "center", font_size = 12)
#print(both_rows[1:5])
```

## References
Hall, J. D., Palsson, C., and Price, J. (2018). Is uber a substitute or complement for public
transit? Journal of Urban Economics, 108:36–50.

Lee, Kyunghee and Jin, Qianran and Animesh, Animesh and Ramaprasad, Jui, Impact of Ride-Hailing Services on Transportation Mode Choices: Evidence from Traffic and Transit Ridership (October 31, 2019). Forthcoming at MIS Quarterly, Available at SSRN: https://ssrn.com/abstract=3244207 or http://dx.doi.org/10.2139/ssrn.3244207

Zhao, W. (2019). The Long Run Effects of Uber on Public Transit, Congestion, Sprawl, and the Environment.
